---
title: "BST244 HW2"
author: "Jonathan Luu"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(tidyverse)
library(survminer)
```

## Problem 1. 

For this homework we will try to estimate survival functions for the time to the first 
recurrent event (variable Recurrence time) in both treatment arms in the bladder recurrence cancer 
study. Please also use the first recurrence time as the survival time (variable Recurrence time), and 
the  corresponding  recurrence  event  as  the  survival  status  (variable  Event,  0  =  censored,  1  = 
recurrence).  Using  the  same  random  90%-sample  subset  you  used  for  last  homework,  for  each 
treatment arm:

```{r, comment=NA}
dat <- read.csv("bladdercancer.csv")

# Get first recurrence
first_dat <- dat %>% group_by(id) %>% slice(1)

# Sample 90% for each treatment group
second_dat <- first_dat %>% group_by(Treatment.group) %>% slice_sample(prop=0.9)
placebo_dat <- second_dat %>% filter(Treatment.group == "1")
treatment_dat <- second_dat %>% filter(Treatment.group == "2")

# Get ids of both treatment arms
placebo_dat$id
treatment_dat$id
```

\newpage
```{r}
# KM Curve
km_curve = survfit(Surv(Recuerrence.time, Event) ~ Treatment.group, 
                   type="kaplan-meier", data = second_dat)
ggsurvplot(km_curve,  xlab="Days", ylab="Survival Probability", 
           title = "Kaplan-Meier Survival Curve", conf.int=T)
```

\newpage
```{r}
# NA Curve
na_curve = survfit(Surv(Recuerrence.time, Event) ~ Treatment.group, 
                   type="fleming-harrington", data = second_dat)
ggsurvplot(na_curve, xlab="Days", ylab="Survival Probability", 
           title = "Fleming-Harrington Survival Curve", conf.int = T)
```

Looks like the confidence intervals plotted for both curves are transformed log intervals.

```{r}
# RMST for both methods
print(km_curve, rmean = 60)
print(na_curve, rmean = 60)
```

## Problem 2.

Using at least two weighting schemes, test for the null hypothesis that the control and 
treatment arms have the same survival function. 

```{r}
# Log-rank test (weight = 1)
fit <- survfit(Surv(Recuerrence.time, Event) ~ Treatment.group, data = second_dat)
ggsurvplot(fit, pval = TRUE, pval.method = TRUE)

# Gehan-Breslow (generalized Wilcoxon) 
ggsurvplot(fit, pval = TRUE, pval.method = TRUE, surv.median.line = "hv",
           log.rank.weights = "n")
```

Both tests show that there is not a significant difference in mean survival time between the two treatment groups. 



